### 2016-02-09 Manuscript/BiologyLetters/response.tex

\textbf{RE: Decision on Manuscript ID RSBL-2015-1003}\\
\bigskip
Dear Surayya Johar,\\
\bigskip
We are very grateful to the four referees for their helpful and constructive comments.s
We have taken all of their comments on board, and respond to their points below. To improve the clarity of this document, we dealt with each comment in the order they appeared in your ``Decision on Manuscript'' email and we have included the reviewers comments in blue.

%%%%%%%%%%%%%%%
% Reviewer 1
%%%%%%%%%%%%%%%

\section{Reviewer 1:}
\begin{enumerate}
\item{\textcolor{blue}{In ESM 1, p. 4, the authors explain their methods of collecting morphological data from the literature. Why select only the first 10 results per search term? How do you know that relevant papers wouldn't come up later in the search? I need evidence to support this cutoff; as it stands, I'm concerned that you're missing some relevant material.}}
% NC: It's not just the first 10 results per search term is it? I thought you went through the first 10 pages of results? Clarify this. TG: No I collected the 20 first results. It's mentioned down there.

As shown in ESM1 Figure 1, the number of OTUs with morphological cladistic data extracted from the Google Scholar searches seems to plateau really quickly (i.e. before the first 100 matches).
Additionally, as Graeme Lloyd mentions, ``older studies tend to be smaller (due to computing limitations) and to commonly reoccur anyway due to data set reuse'' (Reviewer 4, comment 1).
Following these two observations, we think that including the 20 most recent results % NC: Is that what you did? TG: yep.
 for each order (and higher taxonomic level) was a good balance between the time spent looking for papers, the number of OTUs extracted and the repeatability of the study.
We are aware however that we might have missed some specific studies (see Reviewer 2 comment 2) but are confident that our protocol gives a good overall picture of the data availability and that the three repositories used before the Google Scholar search (MorphoBank, Graeme Lloyd's and Ross Mounce's) contained most of the available data.
In fact, as mentioned in ESM 1, part 1 ``among the 660 papers [collected from the Google Scholar search], only 50 contained a total of 425 extra living OTUs''.

@@

\item{\textcolor{blue}{Similarly, I find the 2010 date cutoff troubling, because I'm not sure that the argument that these recent studies contain a fraction of the previously published characters and OTUs allows you to make a comprehensive statement about the availability of character data. What about the fraction that isn't included? This is a question that needs to be better addressed before you will convince readers that this is really all the data that is out there.}}

We chose this cutoff date to filter our analysis towards accessible data only (see Reviewer 2 comment 2).
Additionally, we observed that in newer studies (i.e. post 2010), living taxa with anatomical data were rarely added an usually reused from former studies (with sometimes the addition of extra characters).
We modified the sentence in the ESM1 part 1:

``For example in primates the character p7 coded first by [1] is reused with the same living species in [2], [3], [4], [5], [5], [6], [7], [8], [9], [10], [11], [12] and [13].''

\noindent to: ``For example, the six living primates used in [1] (\textit{Aotus trivirgatus, Galago demidoff, Lemur catta, Microcebus murinus, Nycticebus coucang and Saimiri sciureus}) and their associated characters are reused along with more living species and characters in [2], [3], [4], [5], [5], [6], [7], [8], [9], [10], [11], [12] and [13].''.

@@

\item{\textcolor{blue}{P3, L78: ``since'' should be ``because.'' Since should be used for time, not causation.}}
We replaced ``since'' with ``because'' throughout the manuscript as appropriate.

\item{\textcolor{blue}{P3, L85: ``not-applicable'' should not be hyphenated.}}
We removed the hyphen.

\item{\textcolor{blue}{P3, L120: How much of a difference does the 25\% or less coverage make to the capacity of Total Evidence phylogenetic analyses to resolve relationships? Does your simulation study have a way of quantifying or describing the probability of getting the ``right'' answer with this little data? It would also help if you could show in your data table what the clades were that had better than 75\% coverage. It's not immediately evident from reading Table 1.}}

The straight relationship between the living species coverage and the Total Evidence phylogeny topology is hard to quantify precisely and depends on both the metric used for measuring topological differences and the method used to infer the tree.
Our simulation showed that topology tend to get closer to a random one as morphological data is missing but that in the case of the missing living species, the other parameters we tested (missing data in the fossil record and number of morphological characters) add no significant effect (i.e. when a certain amount of living taxa are missing, the other parameters do not have any extra influence).
In our simulations we tested 5 different level of missing living taxa (0\%, 10\%, 25\%, 50\% and 75\% of missing living taxa) and we observed that the three first levels had no significant negative effect on topology in relation with the amount of missing data in the fossil record and the number of morphological characters but that the fifth level (75\% of missing living taxa, i.e. 25\% coverage) had a strong negative effect.

Also, we highlighted the different levels of coverage in Table 1 as follow: at any taxonomic level, orders with less than 25\% coverage are highlighted in blue (or dark grey in the black and white version) and orders with more than 75\% coverage are highlithed in orange (or light grey in the black and white version).

@@

\end{enumerate}

%%%%%%%%%%%%%%%
% Reviewer 2 (Graham Slater)
%%%%%%%%%%%%%%%

\section{Reviewer 2 (Graham Slater):}
\begin{enumerate}
\item{\textcolor{blue}{The authors state on page 3 lines 82-90 that they avoided matrices with ``few characters'' due to lower probability of overlapping characters with other matrices and so only selected matrices containing >100 characters. First, a lack of overlapping characters doesn’t seem justifiable, at least to my mind, as a reason to exclude a matrix if the ultimate goal is to infer the placement of extinct taxa relative to extant taxa. Antler ramifications, to use the authors' example, are likely to be a very useful character when attempting to resolve relationships among living and extinct artiodactyls, even if they cannot be coded for monotremes. To flip the argument, a matrix of 200 + characters for mysticetes is similarly likely to contain many non-applicable characters from the perspective of talpids. I do understand that the authors are thinking downstream in terms of super matrices for higher level analyses, but smaller matrices are going to be just as relevant for reaching this goal. You don't have much space in the ms I understand, but surely this needs to be addressed?!}}

We removed the analysis with a threshold of OTUs with at least 100 characters and replaced it with the analysis without any minimal number of morphological characters.
We retracted the paragraph about the character threshold analysis and updated the results of figure 1 and table 1 to contain all living OTUs; we removed the redundant ESM 2 and we updated the figures

@@

\section{Reviewer 3 (Peter Wagner):}
\begin{enumerate}
\item{\textcolor{blue}{My biggest criticism is a semantic one: Morphological data $\neq$ cladistic data.}}
We agree with this reviewers semantinc point and changed ``cladistic'' and ``morphological'' data to ``anatomical characters'' throughout the manuscript.

\item{\textcolor{blue}{The real issue here is the availability of phylogenetic character coding for anatomical data.
[...] Thus, the title really should be ``Assessment of available anatomical characters for phylogenetic analysis among living mammals'' or something like that.}}
We changed the title as the reviewer suggested.

\item{\textcolor{blue}{what we really need is to focus on the tooth and skull characters that typify the fossil record.
[...] Still, several papers estimate that the sampling of mammal species (based on cranial characters include teeth) actually is pretty darned good [...].
Thus, we might be able to richly populate FBD and tip-dating analyses with a LOT of extinct species.}}

We totally agree with this comment.
In fact, among the 286 matrices used in this analysis, we found 5228 unique OTUs including 3627 fossil ones which is pretty impressive knowing the amount of work required to collect data for a single specimen!
However, in this study, we focused only on the amount of living taxa (1601 unique OTUs containing 847 species) as a way to solve the problems associated with the lack of living taxa on tree topology described in Guillerme \& Cooper 2016.

@@

\end{enumerate}

\subsection{Reviewer 3 specific comments:}
\begin{enumerate}
\item{\textcolor{blue}{Lines 25-26: ``We suggest that increased morphological data collection efforts for living taxa are needed to produce accurate Total Evidence phylogenies.''
In a way, this is more than ``total evidence.=''.
The initial total evidence paradigm was simply to use every character you could (molecular, morphological, even behavioral).
We now are talking about tip-dating and fossilized birth-death (FBD) approaches (e.g., Heath et al. 2014 PNAS 111:E2957) that want to use all of the available data.
After all, one can do a total evidence analysis with just node-dating techniques rather than with tip-dating, and one can do it without worrying about branch length priors.
However: if we want to use all of the data to calibrate branch lengths for all taxa (fossil and extant), then we need ``total evidence'' to do tip-dating and FBD properly).
As tip-dating is the main theme of this issue, that should be linked to the total evidence explicitly here and elsewhere in the manuscript.}}

\item{\textcolor{blue}{Lines 70-71: ``We downloaded all cladistic matrices containing any living and/or fossil mammal  taxa from three major public databases:''
As a side note and for future reference for the authors, a large number of phylogenetic data sets for mammals (and other metazoans) also are available at the Paleobiology Database.
(These were formerly at my website at the Field Museum, but that site obviously disappeared when I moved to the Smithsonian.)
Because of the nature of the projects for which I was using these datasets, most of them are species-level, but some are genus-level, too.}}

We thank the reviewer for this mention and we will be sure to use this database for future studies.
% NC: If you end up needing to redo the data collection it might be worth checking the PBDB...

@@

\item{\textcolor{blue}{Lines 164-166: ``For example, a Carnivora fossil will be unable to branch in the Herpestidae, and will have more chance to randomly branch within Canidae (Figure 1B).'' 
This seems to be an awkward way to phrase this.
Moreover, this seems like an extension of the long-branch problem.
Basically, the issue really is that we do not have coded skeletal characters for extant herpestids.
Thus, it is not possible for coded skeletal characters from fossil herpestids to connect those fossil herpestids to the extant herpestids.
This is the opposite of Felsenstein's long branch: insofar as skeletal characters are concerned, extant herpestids have zero-length branches.
That in turn means that the fossil herpestids are going to be ``attracted'' to non-zero length branches elsewhere, with long branches in related groups such as canids being particularly good candidates because there is a higher chance of homoplasies between canids and herpesitds on the longer branches.
So, it will not be random: they probably will latch themselves onto the longest reasonably branch among the closest relatives for which anatomical data are present.}}

We do agree with this comment and changed this comment to:
``For example, a Carnivora fossil will be unable to be placed in the Herpestidae clade due to the fact that they have no anatomical characters available (i.e. extant Herpestidae have a zero-length branches for anatomical characters).
Such fossil, however, will have a higher probability to be placed along a branch that contains many anatomical characters (i.e. long branches for anatomical characters, e.g. Canidae or Ursidae (Figure 1B).''

% NC: Yep he's right, modify the sentence to account for this. Also cite a long branch attraction stuff. Check if Pete has written anything - it'd be nice to cite him as he's been so helpful. % TG: I looked for relevant paper on this "short-branch repulsion" but didn't found any that's mildly relevant from Peter...

@@

\end{enumerate}

%%%%%%%%%%%%%%%
% Reviewer 4 (Graeme Lloyd)
%%%%%%%%%%%%%%%

\section{Reviewer 4:}
\begin{enumerate}
\item{\textcolor{blue}{Reviewer 1's most substantial complaints (the year and page number cut offs for google scholar searches) seem at least partially valid.
However, in my experience older studies tend to be smaller (due to computing limitations) and to commonly reoccur anyway due to data set reuse.
The authors also already state that subsequent search results tended to yield no substantial additional data, and by the reviewer's own admission the assembled data seem to look correct.
Thus a recollecting of the data doesn't seem necessary (but see further comments below).}}

We developed description of our data collection protocol in the ESM 1 (see Reviewer 1 comment 2) and underlined the caveats in the discussion (see Reviewer 2 comment 2).

@@

\item{\textcolor{blue}{Reviewer 2 (Graham Slater)'s complaint are - in my view - the most substantial.
I agree that the >100 characters cut off is not as strongly justified as it might be.
However, as the authors have run their protocol without this threshold this is easily fixed by at least briefly mentioning how this affects their results in the main text (not relegating it to the ESM where no actual discussion is provided).
From my reading of the two tables only 11 of the orders seem to have below 25\% coverage at the species-level (not 22 as the authors state in the main text!) when the threshold is used, and this improves slightly to 10 for when the threshold is removed.
As this difference isn't dramatic it seems worth mentioning.
Reviewer 2 also notes that some important studies may be missed.
This seems valid (I am aware of the biases in Ross Mounce's data collection protocol, and obviously my own) and I am sure there are many more data sets that could be sampled.
In particular I am aware of a number of studies that can only be found in books and hence are not easily accessed even if they occur in a google scholar search.
I do not expect the authors to more thoroughly search the literature for all possible cladistic analyses of mammals, however, as this is a major task.
(I should know, it has been a long term project of mine for years.)
Nevertheless, it does deserve mention in the discussion as a potential bias in the results.
I don't think this undermines the study, but rather should be viewed as both a reason for optimism and an acknowledgement of a separate problem (i.e., access to cladistic data) which multiple publications already cover (e.g., see Ross Mounce's work).}}

We removed the character threshold in the whole analysis and specified what we meant by ``data accessibility'' in the discussion (see Reviewer 2 comment 2).

@@

\item{\textcolor{blue}{Reviewer 3 (Pete Wagner)'s major problem - with which I agree - is semantic, and thus easily fixable with a few explicit caveats (e.g., ``by cladistic we mean...'').}}

\item{\textcolor{blue}{One point I would add of my own is that the orders used are obviously very variable in size (species number), so what about total coverage?
This could surely be synthesised from the tables and would be a nice simple way to compare the with and without threshold results.
Similarly, I think Reviewer 3's [We assume Graeme Lloyd is refering to Graham Slater's reviewe (reviewer 2)] %or same as comment before: no names!
 graphs are a nice potential addition to the ESM.}}

We added the data coverage among all mammals in the Table 1 (``Mammalia (Class)'') and added it to the methods, results and discussion.
We also added the Reviewer 2's extra analysis suggestion to the ESM 2 part 1.

@@

\end{enumerate}

\subsection{Reviewer 4 specific comments:}
\begin{enumerate}
\item{\textcolor{blue}{Lines 72-74 - Web addresses should be consistent in dropping of ``http://www.''.}}
We removed the ``http://www.'' part of the URLs.

\item{\textcolor{blue}{Line 104 - ``NTI 15'' should probably be ``NTI [15]''.}}
%TO DO IN THE WORD MANUSCRIPT! 
We fixed this typo.

\item{\textcolor{blue}{Line 104 - It's not completely clear to me how ``distance'' is measured. Is this branch-lengths on a time-scaled tree? Or just a node count?}}
We have clarified this by changing ``observed mean distance between each of $n$ taxa'' to ``observed mean sum of the branch lengths between each of $n$ taxa''.

\item{\textcolor{blue}{Line 117 - It's not clear to me how supraspecific values were calculated if the reference phylogeny (ref 10) is species-level. Are you using the node subtending the higher taxon or something else?}}
We pruned Bininda-Emonds et al's phylogeny for each order in order to include only the species within this order using the the functions:

\texttt{nodeX <- ape::getMRCA(treeAllMammals, listOfSpeciesFromOrderX)}\\
\texttt{treeOrderX <- ape::drop.tip(treeAllMammals, nodeX)}\\

We added the following sentence to in the methods:
``For each analysis our focal taxa were those with available anatomical characters at that taxonomic level and the phylogeny was that of the order pruned from [10]. For each order, we pruned the phylogeny from [10] to include only the species/genus/families from this order.''. p@. lines@@.

\item{\textcolor{blue}{Line 149 - ``most palaeontological studies use the genus as their smallest OTU'' I'm not sure if this is strictly true. It's more that shorter OTU names are commonly used. After all at least one actual specimen has to be coded and in my experience if you read the paper usually only a single species is used. Palaeontologists are just bad at using proper binomials in their matrices (which is an issue, but a different one). In addition ref 20 is actually all species-level OTUs! (I didn't check ref 21.)}}

We modified this statement by stating that ``many palaeontological studies use the genus (or monospecific genera) as their smallest OTU''.

\item{\textcolor{blue}{Line 154 - ``phylogenetically overdispersed'' Purely semantic, but I guess what you mean is that even sampling would be ideal, and this is *measured* using phylogenetic over-dispersion.}}

We changed ``the ideal scenario is for them to be phylogenetically overdispersed'' to ``the ideal scenario is for them to be evenly distributed (as measured by phylogenetic overdispersion)''.

\item{\textcolor{blue}{Line 245 - The caption could be more informative regarding the random and clustered distributions that these figures represent.}}

We changed the caption to: ``Phylogenetic distribution of species with available cladistic data across two orders (A: Primates; B: Carnivora). Blue branches indicate available cladistic data for the species. Available cladistic data is randomly distributed in Primates (A) and clustered in Carnivora (B).''.

\item{\textcolor{blue}{ESM 1 - Incomplete sentence: ``The list of all the 286 download matrices is available on''}}

We fixed this typo. See reviewer 1 specific comment 12.

\item{\textcolor{blue}{ESM 1 - Incomplete sentence: ``All the standardised matrices are available on''}}

We fixed this typo. See reviewer 1 specific comment 12.

\item{\textcolor{blue}{ESM 1 - ``designated as ``fossil'' all OTUs that were present in the Paleobiology database'' The database contains extant taxa too, although these should be marked as such. I guess this doesn't matter given the flow chart in Figure 2, but is something the authors should be aware of. (Obviously more generally living species can have fossil records, or at least plenty of fossils are assigned to extant species.)}}

We do agree with this point (in a similar way: \textit{Thylacinus cynocephalus} is present in Wilson Reader's list but can not be designated as ``living'') and thus added a precision to what we mean by ``fossil'' OTU in the ESM 1: ``i.e. OTUs not present in the two previous data sets [Wilson Reeder and Binida-Emonds]''.

\item{\textcolor{blue}{ESM 1 Figure 2 - How many OTUs were ignored? And based on the (assumed) cursory inspection of the authors what are they? Higher taxa? Specimens? Or just species you can't reconcile with your three data sets?}}

1549 of the 11010 OTUs were ignored.
We added the following sentence to the ESM 1 to explain which OTUs were considered as NAs.
``Non-applicable OTUs were either specimen IDs with no related taxa (e.g. \textit{FMNHPR2081}), abbreviation that were not described in the associated paper (e.g. \textit{Ho.sap.}), non-mammals \textit{stricto-sensu} (e.g. \textit{Sinoconodon}), non standard taxonomic level (e.g. \textit{Spalcotheriids}) invalid taxonomic designation (e.g. \textit{sp\_nov\_1} or \textit{Outgroup}) or typos (e.g. \textit{Hobo sabpiens}).''

### 2016-02-09 Manuscript/BiologyLetters/response.tex

\item{\textcolor{blue}{ESM 2 - ``The following section contains supplementary results to the main body: the available data structure using the NTI and the PD metric'' Abbreviations should be introduced when first used, not later as they are here.}}
We changed this to ``using the Net Relatedness Index (NRI) and the Nearest Taxon Index (NTI) metrics''.

### 2016-02-09 Manuscript/BiologyLetters/Missing_mammals-draft.tex
Updated response (finished)

\subsection{Data availability and distribution}
To assess the availability of anatomical characters for each mammalian order, we calculated the percentage of OTUs with cladistic data at three different taxonomic levels: family, genus and species. %TG: changed cladistic datat to anatomical characters
We consider orders with \textless 25\% of living taxa with available anatomical characters as having low data coverage, and orders with \textgreater 75\% of living taxa with available anatomical characters as having high data coverage.

For each order, we investigated whether the available anatomical characters were (i) randomly distributed, (ii) overdispersed or (iii) clustered, with respect to phylogeny, using two metrics from community phylogenetics: the Nearest Taxon Index (NTI; \cite{webb2002phylogenies}) and the Net Relatedness Index (NRI; \cite{webb2002phylogenies}). 
NTI is most sensitive to clustering or overdispersion near the tips, whereas NRI is more sensitive to it across the whole phylogeny \cite{Cooper2008}. 
Both metrics were calculated using the \texttt{picante} package in R \cite{picante,R}.

@@

\section{Results}
11 out of 28 orders have low coverage (\textless 25\% species with available anatomical characters) and 7 have high coverage (\textgreater 75\% species with available anatomical characters) at the species-level.
At the genus-level, one order have low coverage and 15 have high coverage, and at the family-level, no orders have low coverage and 25 have high coverage (Table\ref{Table_results}).

@@

Only seven orders had significantly clustered data (Cetartiodactyla, Cingulata, Pilosa and Rodentia at the species-level and Carnivora, Chiroptera and Soricomorpha at both species- and genus-level) and none had significantly overdispersed data (Table~\ref{Table_results}).

@@

The number of living mammalian OTUs with no available anatomical characters was surprisingly high at the species-level: only seven out of 28 orders have a high coverage of taxa with available anatomical characters.
This high coverage threshold of 75\% of taxa with available anatomical characters represents the minimum amount of data required before missing data has a significant effect on the topology of Total Evidence trees \cite{GuillermeCooper}.
Beyond this threshold, there is considerable displacement of wildcard taxa and decreased clade conservation \cite{GuillermeCooper}.
Therefore we expect difficulties in placement of fossils at the species-level in most mammalian orders, but fewer issues at higher taxonomic levels.
This point is important from a practical point of view because of the slight discrepancy between neontological and palaeontological species concepts.
While neontological species are described using morphology, genes, distribution etc.; palaeontological species can be based only on morphological, spatial and temporal data (e.g. \cite{ni2013oldest}).
Therefore, many palaeontological studies use the genus (or monospecific genera) as their smallest OTU (e.g. \cite{ni2013oldest,O'Leary08022013}), so data availability at the genus-level in living mammals should be our primary concern when building phylogenies of living and fossil taxa.

When few species have available anatomical characters, the ideal scenario is for them to be evenly distributed (as measured by phylogenetic overdispersion) to maximize the possibilities of a fossil branching in the right clade.
The second best scenario is that species with available anatomical characters are randomly distributed across the phylogenies. 
Here we expect no special bias in the placement of fossils \cite{GuillermeCooper}, it is therefore encouraging that for most orders, species with available anatomical characters were randomly distributed across the phylogeny.
The worst case scenario for fossil placement is that species with available anatomical characters are phylogenetically clustered. 
Then we expect two major biases to occur: first, fossils will not be able to branch within a clade containing no data, and second, fossils will have higher probability of branching within the most sampled clade by chance. 
Our results suggest that this may be problematic at the genus-level in Carnivora, Chiroptera and Soricomorpha. 
For example, a Carnivora fossil will be unable to be placed in the Herpestidae clade% rev 1.2.7
, and will have more chance to randomly branch within Canidae or Ursidae (Figure \ref{Figure_example_coverage}B).

>>>>

\subsection{Data availability and distribution}
To assess the availability of anatomical characters for each mammalian order and for Mammalia as a class, we calculated the percentage of OTUs with cladistic data at three different taxonomic levels: family, genus and species. %TG: changed cladistic datat to anatomical characters
We consider orders with \textless 25\% of living taxa with available anatomical characters as having low data coverage, and orders with \textgreater 75\% of living taxa with available anatomical characters as having high data coverage.

For each order and for Mammalia as a class, we investigated whether the available anatomical characters were (i) randomly distributed, (ii) overdispersed or (iii) clustered, with respect to phylogeny, using two metrics from community phylogenetics: the Nearest Taxon Index (NTI; \cite{webb2002phylogenies}) and the Net Relatedness Index (NRI; \cite{webb2002phylogenies}). 
NTI is most sensitive to clustering or overdispersion near the tips, whereas NRI is more sensitive to it across the whole phylogeny \cite{Cooper2008}. 
Both metrics were calculated using the \texttt{picante} package in R \cite{picante,R}.

@@

\section{Results}
Among all Mammalia, the species coverage was low (\textless 25\% species with available anatomical characters) but the family coverage was high (\textgreater 75\% species with available anatomical characters).
For each order, 11 out of 28 have low coverage and 7 have high coverage at the species-level.
At the genus-level, one order have low coverage and 15 have high coverage, and at the family-level, no orders have low coverage and 25 have high coverage (Table\ref{Table_results}).

@@

Among all Mammalia, the data was significantly cluster when using the NTI at the species and genus level.
For each order, only seven had significantly clustered data (Cetartiodactyla, Cingulata, Pilosa and Rodentia at the species-level and Carnivora, Chiroptera and Soricomorpha at both species- and genus-level) and none had significantly overdispersed data (Table~\ref{Table_results}).

@@

The number of living mammalian OTUs with available anatomical characters was surprisingly low at the species-level: only 17\%.
Only seven out of 28 orders have a high coverage of taxa with available anatomical characters.
This high coverage threshold of 75\% of taxa with available anatomical characters represents the minimum amount of data required before missing data has a significant effect on the topology of Total Evidence trees \cite{GuillermeCooper}.
Beyond this threshold, there is considerable displacement of wildcard taxa and decreased clade conservation \cite{GuillermeCooper}.
Therefore we expect difficulties in placement of fossils at the species-level in most mammalian orders, but fewer issues at higher taxonomic levels.
This point is important from a practical point of view because of the slight discrepancy between neontological and palaeontological species concepts.
While neontological species are described using morphology, genes, distribution etc.; palaeontological species can be based only on morphological, spatial and temporal data (e.g. \cite{ni2013oldest}).
Therefore, many palaeontological studies use the genus (or monospecific genera) as their smallest OTU (e.g. \cite{ni2013oldest,O'Leary08022013}), so data availability at the genus-level in living mammals should be our primary concern when building phylogenies of living and fossil taxa.

When few species have available anatomical characters, the ideal scenario is for them to be evenly distributed (as measured by phylogenetic overdispersion) to maximize the possibilities of a fossil branching in the right clade.
The second best scenario is that species with available anatomical characters are randomly distributed across the phylogenies. 
Here we expect no special bias in the placement of fossils \cite{GuillermeCooper}, it is therefore encouraging that for most orders, species with available anatomical characters were randomly distributed across the phylogeny.
The worst case scenario for fossil placement is that species with available anatomical characters are phylogenetically clustered. 
Then we expect two major biases to occur: first, fossils will not be able to branch within a clade containing no data, and second, fossils will have higher probability of branching within the most sampled clade by chance. 
Our results suggest that this may be problematic at the genus-level in Carnivora, Chiroptera and Soricomorpha. 
For example, a Carnivora fossil will be unable to be placed in the Herpestidae clade due to the fact that they have no anatomical characters available (i.e. extant Herpestidae have a zero-length branches for anatomical characters).
Such fossil, however, will have a higher probability to be placed along a branch that contains many anatomical characters (i.e. long branches for anatomical characters, e.g. Canidae or Ursidae (Figure \ref{Figure_example_coverage}B).

### 2016-02-09 Manuscript/BiologyLetters/Supplementary/ESM_1.tex
Updated email
Changed anatomical characters semantics
Updated response (finished)

\subsubsection{Ross Mounce}
We accessed Ross Mounce's GitHub repository (\url{github.com/rossmounce/cladistic-data}) on the 11th of June 2015 and downloaded all 601 matrix.
We then ran a shell script to select only the matrices that had any text element that match with one of the search terms.
% LINK WHICH SHELL SCRIPT
To make the matrix selection more thorough, we ignored the keywords case as well as the latin suffix (\textit{ia}, \textit{ata}, \textit{ea}, and \textit{a}).

@@

@@

\subsection{Selecting the living OTUs}
%To select the lowest valid taxonomic level for each OTU, we standardised their taxonomy by correcting species names so they matched standard taxonomic nomenclature (e.g., \textit{H. sapiens} was transformed to \textit{Homo sapiens}).
We designated as ``living'' all OTUs that were either present in the phylogeny of \cite{bininda-emondsthe2007} or the taxonomy of \cite{wilson2005mammal}, and designated as ``fossil'' all OTUs that were present in the Paleobiology database (\url{https://paleobiodb.org/}).
For OTUs that did not appear in these three sources, we first decomposed the name (i.e. \textit{Homo sapiens} became \textit{Homo} and \textit{sapiens}) and tried to match the first element with a higher taxonomic level (family, genus etc.).
Any OTUs that still had no matches in the sources above were designated as non-applicable (NA; Figure ~\ref{Supp_figure_Taxonomic_algorithm}).

>>>>

\subsubsection{Ross Mounce}
We accessed Ross Mounce's GitHub repository (\url{github.com/rossmounce/cladistic-data}) on the 11th of June 2015 and downloaded all 601 matrix.
We then ran a shell script to select only the matrices that had any text element that match with one of the search terms (\url{https://github.com/TGuillerme/Missing_living_mammals/blob/master/Functions/select.files.sh}).
To make the matrix selection more thorough, we ignored the keywords case as well as the latin suffix (\textit{ia}, \textit{ata}, \textit{ea}, and \textit{a}).

@@

For example, the six living primates used in \cite{ross1998phylogenetic} (\textit{Aotus trivirgatus, Galago demidoff, Lemur catta, Microcebus murinus, Nycticebus coucang and Saimiri sciureus}) and their associated characters are reused along with more living species and characters in \cite{seiffert2003fossil}, \cite{marivaux2005anthropoid}, \cite{seiffert2005basal}, \cite{bloch2007new}, \cite{bloch2007new}, \cite{kay2008anatomy}, \cite{silcox2008biogeographic}, \cite{seiffert2009convergent}, \cite{tabuce2009anthropoid}, \cite{boyer2010astragalar}, \cite{seiffert2010fossil}, \cite{marivaux2013djebelemur} and \cite{ni2013oldest}.

@@

\subsection{Selecting the living OTUs}
%To select the lowest valid taxonomic level for each OTU, we standardised their taxonomy by correcting species names so they matched standard taxonomic nomenclature (e.g., \textit{H. sapiens} was transformed to \textit{Homo sapiens}).
We designated as ``living'' all OTUs that were either present in the phylogeny of \cite{bininda-emondsthe2007} or the taxonomy of \cite{wilson2005mammal}, and designated as ``fossil'' (i.e. OTUs not present in the two previous data sets) all OTUs that were present in the Paleobiology database (\url{https://paleobiodb.org/}).
For OTUs that did not appear in these three sources, we first decomposed the name (i.e. \textit{Homo sapiens} became \textit{Homo} and \textit{sapiens}) and tried to match the first element with a higher taxonomic level (family, genus etc.).
Any OTUs that still had no matches in the sources above were designated as non-applicable (NA; Figure ~\ref{Supp_figure_Taxonomic_algorithm}).
Non-applicable OTUs were either specimen IDs with no related taxa (e.g. \textit{FMNHPR2081}), abbreviation that were not described in the associated paper (e.g. \textit{Ho.sap.}), non-mammals \textit{stricto-sensu} (e.g. \textit{Sinoconodon}), non standard taxonomic level (e.g. \textit{Spalcotheriids}) invalid taxonomic designation (e.g. \textit{sp\_nov\_1} or \textit{Outgroup}) or typos (e.g. \textit{Hobo sabpiens}).
